---
title: "homework 13"
---

Using a dataset of your choice, ask two separate questions about the data and construct an appropriate model and figure to test and visualize each one.

Using the starwars dataset:

### Load libraries:
```{r}

library(tidyverse)
library(lubridate)
library(lme4)
library(car)

```

### Load in Dataset:
```{r}
sw <- starwars
```

## Question 1: Can you predict height based off of sex and mass, while accounting for species?

### Creating the model:
```{r}
# loading in starwars dataset, filtering out an outlier (Jabba the Hutt)
cleanstar <- starwars |>
  filter(mass<1000)

# creating model for height as a factor of sex and mass, accounting for species
star_gamma <- lmer(
  height ~  sex * mass + (1 | species),
  data = cleanstar)
anovagamma <- Anova(star_gamma)
anovagamma
# This analysis included a non-Gaussian mixed model with random effects
```

### Plotting:
```{r}
height_mass_sex_spp <- ggplot(data = cleanstar) +
  aes(x = height, y = mass) +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_point(mapping = aes(
    color = factor(species),
    shape = sex
  )) + geom_smooth()

height_mass_sex_spp

```

### Explanation:

In this model, I used sex and mass to predict the height of individuals, while accounting for a random effect variable of species. I used species to  While mass was a significant predictor (<2e-16), sex was not. I believe that sex is not significant due specifically because I accounted for species, because there are many more males in the starwars dataset, and so not all species have both females and males in them. 


## Question 2: Does gender (feminine or not feminine) predict height in starwars?

### Creating model:
```{r}
# adding a column to create a binary for if an individual is or isn't feminine
cleanstar <- cleanstar |> mutate(
    gender_tf = gender == "feminine"
) |> na.omit()

# cat y, cont x

star_cat_cont <- glm(
  gender_tf ~ height,
  data = cleanstar,
  family = binomial(link="logit")
)
Anova(star_cat_cont)

```

### Plotting model:
```{r}

star_cat_cont_plot <- ggplot(
    data = cleanstar,
    mapping = aes(
        x = gender_tf,
        y = mass
    )
) + geom_point() + geom_smooth(
    method = "glm",
    method.args = list(family = "binomial")
)

star_cat_cont_plot

```

### Explanation: 

In this model, I used if a character was or wasn't feminine to predict height. Because this was a categorical Y x continuous X plot, I used a binomial GLM model. In this model, this metric of gender was not a significant predictor of height (=0.2565). 